Learning and comparing functional connectomes across subjects

### 2013-01-01 paper.tex
iter

Horwitz \emph{el al.}\ \cite{horwitz1995} claimed almost 20 years ago
that ``the crucial concept needed for network analysis is covariance''.
In our eyes, this still holds today. Estimation functional connectomes
relies largely on fitting covariance models. Their comparison requires
understanding how these covariances vary and finding metrics to capture
this variability. Additionally, a secret ingredient to all statistical
steps is a good choice of confounds.
% XXX: more practical advice?

Methodological challenges to functional-connectome-based group studies
arise from the dimensionality and the of variability the connectome. With
the current tools, inter-subject comparison of connectomes comprising
many nodes is limited by the difficulty of estimating high-dimensional
covariance matrices and the loss of statistical power due to multiple
comparisons. Better algorithms integrating powerful a priori is required
to push the limits of covariance estimation. Better characterization of
inter-subject variability of connectomes \cite{kelly2012} will help
choosing parameterizations and invariants to avoid testing each edge for
a difference, as it inevitably leads to a needle in a haystack problem.

Reviewing methodological options to learn and compare connectomes
highlights that there is currently no unique solution, but a spectrum of
related methods and analysis strategies. More empirical results are
required to guide the choices, however this diversity is probably
unavoidable: a diffuse disease like schizophrenia will not lead to the
same connectome modifications as a focal lesion. In statistical learning,
``no free lunch'' theorems \cite{wolpert1996} tell us that no strategy
can perform uniformly better in all situations. In practice, the key to a
successful analysis is to understand well the assumptions and
interpretation of each option, in order to match the method to the
question.

>>>>

Horwitz \emph{el al.}\ \cite{horwitz1995} claimed almost 20 years ago
that ``the crucial concept needed for network analysis is covariance''.
In our eyes, this still holds today. Estimation functional connectomes
relies largely on fitting covariance models. Their comparison requires
understanding how these covariances vary and finding metrics to capture
this variability. The additional secret ingredient may be a good choice
of confounds in all statistical steps. A good choice of a small number of
relevant regions facilitates connectome comparison. However, such a
choice cannot yet be fully factored out via methods and must rely on
neuroscientific expertise.

Methodological challenges to functional-connectome-based group studies
arise from the dimensionality and the of variability the connectome. With
the current tools, inter-subject comparison of connectomes comprising
many nodes is limited by the difficulty of estimating high-dimensional
covariance matrices and the loss of statistical power due to multiple
comparisons. Better algorithms integrating powerful a priori are required
to push the limits of covariance estimation. Better characterization of
inter-subject variability of connectomes \cite{kelly2012} will help
choosing parameterizations and invariants to avoid testing each edge for
a difference, as this strategy inevitably leads to a needle in a haystack
problem.

Reviewing methodological options to learn and compare connectomes
highlights that there is currently no unique solution, but a spectrum of
related methods and analysis strategies. More empirical results are
required to guide the choices. However this diversity is probably
unavoidable: a diffuse disease like schizophrenia will not lead to the
same connectome modifications as a focal lesion. In statistical learning,
``no free lunch'' theorems \cite{wolpert1996} tell us that no strategy
can perform uniformly better in all situations. In practice, the key to a
successful analysis is to understand well the assumptions and
interpretation of each option, in order to match the method to the
question.
